---
title: "Assignment7"
author: "Ruinan Wang"
date: "17/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(Stat2Data)
library(ggplot2)
library(dplyr)
library(PairedData)
library(purrr)
```

## Student’s t-confidence intervals

```{r}
adelie_flippers <- penguins %>% filter(species=="Adelie") %>% pull(flipper_length_mm) %>% na.omit()
alpha<-0.05
sample_size<-length(adelie_flippers)
sample_mean<-mean(adelie_flippers)
sample_sd<-sd(adelie_flippers)
t<-qt(1-alpha/2,df=sample_size-1)
confidence_interval_l<-sample_mean-t*sample_sd/sqrt(sample_size)
confidence_interval_u<-sample_mean+t*sample_sd/sqrt(sample_size)
confidence_interval<-c(confidence_interval_l,confidence_interval_u)
confidence_interval


data("Hawks")
head(Hawks)
red_tailed_Hawks_weights <- Hawks %>% filter(Species == "RT") %>% pull(Weight)%>% na.omit()

alpha <- 0.01
sample_size <- length(red_tailed_Hawks_weights)
sample_mean <- mean(red_tailed_Hawks_weights)
sample_sd <- sd(red_tailed_Hawks_weights)

t <- qt(1-alpha/2,df=sample_size-1)
l <- mean(sample_mean-t*sample_sd/sqrt(sample_size))
u <- mean(sample_mean+t*sample_sd/sqrt(sample_size))
confidence_interval <- c(l,u)
confidence_interval

tibble(red_tailed_Hawks_weights) %>% ggplot(aes(x=red_tailed_Hawks_weights))+geom_density()+theme_bw()+labs(x="Weight (grams)", y="density")


tibble(red_tailed_Hawks_weights) %>% ggplot(aes(sample=red_tailed_Hawks_weights))+stat_qq()+stat_qq_line(color="blue")+theme_bw()
```

##### Before deriving the confidence intervals based on T distribution, we suppose the sample data could be subject to normal distribution


## 2 One sample t-test

```{r}
bill_adelie <- penguins %>% filter(species == "Adelie") %>% pull(bill_length_mm) %>% na.omit()
alpha<- 0.01
sample_size <- length(bill_adelie)
sample_mean <- mean(bill_adelie)
sample_sd <- sd(bill_adelie)
t<-qt(1-alpha/2,df=sample_size-1)
l<-sample_mean-t*sample_sd/sqrt(sample_size)
u<-sample_mean+t*sample_sd/sqrt(sample_size)

confidence_interval <- c(l,u)
confidence_interval

t.test(bill_adelie,conf.level=0.99,mu=40)
```
##### (The sample data is supposed to be subject to normal distribution)


## 3 Implementing a one-sample t-test

```{r}
two_sides_ones_sample_Ttest <- function (X, mu0) {
  sample_mean <- mean(X)
  sample_sd <- sd(X)
  sample_size <- length(X)
  test_statistic <- (sample_mean-mu0)/(sample_sd/sqrt(sample_size))
  p_value <- 2*(1-pt(abs(test_statistic), df=sample_size-1))
  return (p_value)
}
two_sides_ones_sample_Ttest(bill_adelie, 40)
```


## 4 The paired t-test
```{r}
data("Barley")
Glabron_yields <- Barley %>% pull(Glabron) %>% na.omit()
Velvet_yields <- Barley %>% pull(Velvet) %>% na.omit()
t.test(Glabron_yields,Velvet_yields)
```

##### Because the significance level is 0.01, but p-value is 0.04101. So, we fail to reject the null hypothesis of $H_0$  (There is difference for these two types of barley on yields.)

```{r}
diffs <-  Glabron_yields - Velvet_yields
effect_size <- mean(diffs)/sd(diffs)
effect_size
```

##### one sample value should be subject to normal distribution

```{r}
tibble(diffs) %>% ggplot(aes(x=diffs))+geom_density()+theme_bw()+labs(x="Yields diffs", y="density")


tibble(diffs) %>% ggplot(aes(sample=diffs))+stat_qq()+stat_qq_line(color="blue")+theme_bw()

length(diffs)
```

this case isn't subject to normal distribution (Severe skewed distribution). 




### 5 Investing Coverage for Student's t intervals

```{r}
student_t_confidence_interval<-function(sample,confidence_level){
  sample<-sample[!is.na(sample)] # remove any missing values
  n<-length(sample) # compute sample size
  mu_est<-mean(sample) # compute sample mean
  sig_est<-sd(sample) # compute sample sd
  alpha = 1-confidence_level # alpha from gamma
  t<-qt(1-alpha/2,df=n-1) # get student t quantile
  l=mu_est-(t/sqrt(n))*sig_est # lower
  u=mu_est+(t/sqrt(n))*sig_est # upper
  return(c(l,u))
}

num_trials<-100000
sample_size<-30
mu_0<-1
sigma_0<-3
alpha<-0.05
set.seed(0) # set random seed for reproducibility
single_alpha_coverage_simulation_df<-data.frame(trial=seq(num_trials))%>%
  mutate(sample=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_0,sd=sigma_0)))%>% # generate random Gaussian samples
  mutate(ci_interval=map(.x=sample,.f=~student_t_confidence_interval(.x,1-alpha)))%>% # generate confidence intervals
  mutate(cover=map_lgl(.x=ci_interval,.f=~((min(.x)<=mu_0)&(max(.x)>=mu_0))))%>% # check if interval covers mu_0
  mutate(ci_length=map_dbl(.x=ci_interval,.f=~(max(.x)-min(.x)))) # compute interval length
single_alpha_coverage_simulation_df%>%
  pull(cover)%>%
  mean() # estimate of coverage probability




```