---
title: "assginment6"
author: "Ruinan Wang"
date: "10/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 A Gaussian model for Red tailed hawks
```{r}
library(Stat2Data)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
data("Hawks")

RedTailedDf <- Hawks %>%
  filter(Species=="RT") %>%
  select(Weight,Tail,Wing)

head(RedTailedDf)

kernel_tail <- RedTailedDf%>%
  pull(Tail)

n <- length(kernel_tail)
mu_mle_tail <- mean(RedTailedDf$Tail,na.rm = 1)
sigma_mle_tail <- sd(RedTailedDf$Tail, na.rm = 1)*sqrt((n-1)/n)

mu_mle_tail

sigma_mle_tail

colors <- c("MLE density" = "red", "Kernel density"="blue")

Tails <- seq(mu_mle_tail-3*sigma_mle_tail,mu_mle_tail+3*sigma_mle_tail, by=sigma_mle_tail*0.001)

ggplot()+geom_line(data=data.frame(Tail=Tails,Density=dnorm(Tails,mu_mle_tail,sigma_mle_tail)),aes(x=Tail,y=Density, color = "MLE density")) + geom_density(data=tibble(kernel_tail),aes(x=kernel_tail,color="Kernel density"))+labs(y="Density function (mm)", color="Estimator")+theme_bw()+scale_color_manual(values=colors)
```


## 2 Location estimators with Gaussian data
```{r}
set.seed(0)
num_trials_per_sample_size<-100
min_sample_size<-5
max_sample_size<-1000
sample_size_inc<-5
mu_0<-1
sigma_0<-3

simulation_df<-crossing(trial=seq(num_trials_per_sample_size), sample_size=seq(min_sample_size, max_sample_size,sample_size_inc))%>% # create data frame of all pairs of sample_size and trial
  mutate(simulation=pmap(.l=list(trial,sample_size),.f=~rnorm(.y,mean=mu_0,sd=sigma_0)))%>% # simulate sequences of Gaussian random variables
  mutate(sample_md=map_dbl(.x=simulation,.f=median))%>% # compute the sample medians
  mutate(sample_mean = map_dbl(.x=simulation, .f=mean))%>%
  group_by(sample_size)%>%
  summarise(msq_error_md=mean((sample_md-mu_0)^2), msq_error_mean1=mean((sample_mean-mu_0)^2),msq_error_mean2=(mean(sample_mean)-mu_0)^2+(mean(sample_mean-mean(sample_mean)))^2)

head(simulation_df)

simulation_df <- simulation_df%>%
  pivot_longer(cols=-sample_size,names_to = "Estimator", values_to="value")

simulation_df$Estimator <-factor(simulation_df$Estimator,levels = c('msq_error_mean1','msq_error_mean2','msq_error_md'),labels=c("Mean1","Mean2","Meidan"))

ggplot(data=simulation_df, aes(x=sample_size,y=value,color = Estimator, linetype=Estimator))+geom_point()+geom_smooth() +theme_bw() + labs(x="Sample size", y= "Mean Square Error")+xlim(5,500)+ylim(0,0.35)
```

## 3  Unbiased estimation of the population variance


### First, let's suppose we have a $X$ with a Gaussian Distribution ~ $N (1,3)$, here are the simulation study
```{r}

simulation_df <- crossing (trial=seq(num_trials_per_sample_size),sample_size=seq(min_sample_size,max_sample_size,sample_size_inc)) %>%
  mutate(simulation=pmap(.l=list(trial,sample_size),.f=~rnorm(.y,mean=mu_0,sd=sigma_0)))%>%
  mutate(sample_variance_MLE=map_dbl(.x=simulation,.f=~var(.x)*(n-1)/n)) %>%
  mutate(sample_variance_U=map_dbl(.x=simulation,.f=var))%>%
  group_by(sample_size)%>%
  summarise(bias_variance_MLE= (mean(sample_variance_MLE)-sigma_0^2), bias_variance_U= (mean(sample_variance_U)-sigma_0^2))

simulation_df

simulation_df <- simulation_df%>%
  pivot_longer(cols=-sample_size,names_to = "Estimator", values_to="value")

simulation_df$Estimator <-factor(simulation_df$Estimator,levels = c('bias_variance_MLE','bias_variance_U'),labels=c("Variance MLE","Variance U"))

ggplot(data=simulation_df, aes(x=sample_size, y=value, color=Estimator, linetype=Estimator)) +geom_smooth(se=FALSE)+theme_bw()+labs(x="Sample Size", y="Bias Value")+ylim(-0.2,0.2) + geom_hline(aes(yintercept=0),color="blue")

```

As we can see from the plot, the bias for variance U is more closed to 0, which means Variance U is more closed to the unbiased estimator.



## 5 Maximum likelihood estimation with the Poisson distribution


```{r}

```


