---
title: "Assignment2"
author: "Ruinan Wang"
date: "06/10/2021"
output: html_document
---

```{r setup, include=FALSE}
options(warn = -1)
knitr::opts_chunk$set(echo = TRUE)
library(Stat2Data)
library(tidyverse)
```

## 1. Types of variables
### check the rows number and columns number of hawksSmall
```{r }
data("Hawks")
hawksSmall<-drop_na(select (Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))
dim(hawksSmall)
```

## Robustness of location estimator


```{r echo = TRUE}
hal <- Hawks$Hallux
hal<- hal[!is.na(hal)]
outlier_val<-100
num_outliers <-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers)
)
num_outliers_vect <-seq(0,1000)
means_vect<-c()
for (num_outliers in num_outliers_vect){
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
means_vect<- c(means_vect,mean(corrupted_hal))
}

num_outliers_vect<-seq(0,1000)
median_vect<-c()

for(num_outliers in num_outliers_vect){
corrupted_hal <-c(hal,rep(outlier_val,times=num_outliers))
median_vect<-c(median_vect, median(corrupted_hal))
}

t_means_vect<-c()
for(num_outliers in num_outliers_vect){
corrupted_hal<c(hal,rep(outlier_val,times=num_outliers))
t_means_vect<- c(t_means_vect,mean(corrupted_hal,na.rm=1,trim=0.1))
}

df_means_medians<- data.frame(num_outliers=num_outliers_vect, mean=means_vect, t_mean= t_means_vect, median=median_vect)
df_means_medians%>%
pivot_longer(!num_outliers, names_to = "Estimator", values_to ="value")%>%
ggplot(aes(x=num_outliers,color=Estimator,linetype=Estimator, y=value)) + geom_line()+xlab("Number of outliners")
```

## Box plots and outliers
```{r echo=TRUE}
Hawks%>%
  ggplot(aes(x=Species,y=Weight,na.rm=1)) + geom_boxplot()
```


### Computer the number of outlier for the three samples of hawk weights broken down by species.
```{r echo=TRUE}
outliers <- function(x){
  q25<-quantile(x,0.25,na.rm = 1)
  q75<-quantile(x,0.75,na.rm = 1)
  iq_range <- q75-q25
  return( x[((x>q75+1.5*iq_range)|(x<q25-1.5*iq_range))&!is.na(x)])
}
num_outliers <- function (x) {return(length(outliers(x)))}

Hawks%>%
  group_by(Species)%>%
  summarise(num_outliers_weight=num_outliers(Weight))
```


### Covariance and correlation under linear transformations

(1)Covariance between $\widetilde{X_i}$ and $\widetilde{Y_i}$ is
$$
  \sum\widetilde{X},\widetilde{Y} = (ac)\sum{X},{Y}
$$

(2) Correlation between $\widetilde{X_i}$ and $\widetilde{Y_i}$ is

$$
\rho{\widetilde{X},{\widetilde{Y}}} = \frac {\rho{X,Y}}{|a||c|}
$$

