---
title: "Assignment"
author: "Ruinan Wang"
date: "2021/10/12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(purrr)
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(palmerpenguins)
```

# Assignment

## 1 Random experienments, events and sample spaces
  
### **Definition:**  
  **Random experiment** is a procedure (real or imagined) which:  
       (a) has a well-defined set of possible outcomes;  
       (b) could (at least in principle) be repeated arbitrarily many times  
  **Event** is a set (i.e a collection) of possible outcomes. (it implies the sample space)  
  **sample space** is the set of all possible outcomes of interest for a random experiment.  
  
      
+ **What is the random experiment in your example?**  
   The football game between Manchester United and Chelsea.  
+ **What are the possible outcomes in your example?**  
   Manchester United wins, Chelsea wins, the two teams draws, the game is interrupted due to unexpected events.  
+ **What are the events in your example?**  
   {Manchester United wins,the two teams draws},{Chelsea wins, Manchester United wins},{Chelsea wins},{the game is interrupted due to unexpected events}.....  
+ **What is the sample space in your example?**  
   {Manchester United wins, Chelsea wins, the two teams draws, the game is interrupted due to unexpected events}   



## 2 Tidy data and iteration  

### 2.1 Missing data and iteration  

The following function performs imputation by mean. What library do we need to load to run this function?   **library(purrr) ## The purrR library which is contained in the Tidyverse**
```{r echo = TRUE, results='hide'}
impute_by_mean<-function(x){
  mu<-mean(x,na.rm=1) # first compute the mean of x
  
  impute_f<-function(z){ # coordinate-wise imputation
    
    if(is.na(z)) {return(mu)}# if z is na replace with mean
    else {return(z)}# otherwise leave in place
  }
    return(map_dbl(x,impute_f))# apply the map function to impute across vector
  }

```




Create a function called impute_by_median which imputes missing values based on the median of the sample,rather than the mean.

```{r echo=TRUE}
impute_by_median <- function(x){
  value <- median(x,na.rm = 1)
  impute_f<-function(z){
    if (is.na(z)){
      return(value)
    }else{return(z)}
  }
  return(map_dbl(x,impute_f))
}

v<-c(1,2,NA,4)
impute_by_median(v)
```
  
  
Next generate a data frame with two variables x and y.  For our first variable x, we have a sequence(x1,x2,...,xn) where x1=0, xn=  10 and for each i = 1,...,n−1, xi+1=xi+ 0.1. For our second variable y we set yi= 5×i+ 1 for i= 1,...,n. Generate data of this form and place within a data frame called df_xy.

```{r echo=TRUE}
generateY <- function(x){return (5*x+1)}
x <- seq(from = 0, to = 10, by =0.1)
y <- map_dbl(x,generateY)
df_xy <- data.frame(x,y)
head(df_xy, n=5)
```


We will now use map2_dbl()to generate a new data frame with missing data.First create a function sometimes_missing with two variables index and value. The function should return NA if index is divisible by 5 and returns value otherwise.

```{r echo=TRUE}
sometimes_missing <- function (index,value){
  if (index %% 5 ==0){
    return(NA)
  }else{return(value)}
}

sometimes_missing(14,25)
sometimes_missing(15,25)
```

Next generate a new data frame called df_xy_missing with two variables x and y, but some missing data.

```{r echo=TRUE}
df_xy_missing <- df_xy %>%
  mutate(y = map2_dbl(row_number(),y,sometimes_missing))
df_xy_missing%>%head(10)

```



Create a new data frame df_xy_impute with two variables x and y, using compute_by_median.
```{r echo=TRUE}
df_xy_impute <- df_xy_missing %>%
  mutate(y = impute_by_median(y))
head(df_xy_impute,n=10)
```

Plot the original data, the corrupted data and the imputed data together together with a trend line for eachsample.
```{r echo=TRUE}
df_xy <- df_xy %>%
  mutate(source="original")
df_xy_missing<-df_xy_missing%>%
  mutate(source="corrupted")
df_xy_impute<-df_xy_impute%>%
  mutate(source="imputed")
df_combined<-rbind(df_xy,df_xy_missing,df_xy_impute)

ggplot(df_combined,aes(x=x,y=y,color=source))+geom_point()+facet_wrap(~source)+geom_smooth(method="lm")
```

Do the imputed values y give reasonable estimates of the true values y?  
&nbsp;&nbsp; ***Don't know how to explain.***     



## 2.2 Tidying data with pivot functions

```{r}
library(readxl)# load the readxl library
folder_path<-"C:\\UoB\\Statistical Computing and Empirical Methods (SCEM)\\firstRProject\\week3"# set this to the name of the directory containing "HockeyLeague.xlsx"
file_name<-"HockeyLeague.xlsx" # set the file name 
file_path<-paste(folder_path,file_name,sep="\\") # create the file_path
wins_data_frame<-read_excel(file_path,sheet="Wins") # read of a sheet from an XML file

wins_data_frame %>%
  select(1:5)%>%
  head(3)
```

   
   + Is the above data tidy?
   No, Columns correspond to single variables with the same sort of value for each observation. But the competition data every year includes the total number of competitions and wins number.
   
```{r}
wins_tidy <- wins_data_frame %>%
  rename(Team="...1")%>%
  pivot_longer(cols=-Team,names_to="Years",values_to = "Wins_of_Total")%>%
  separate(col="Wins_of_Total",into = c("Wins","Total"),sep = " of ", convert = TRUE)
head(wins_tidy, 5) 
dim(wins_tidy)
  
```

The data frame about complete hockeyleague

```{r}
losses_data_frame<-read_excel(file_path,sheet="Losses") # read of a sheet from an XML file

losses_data_frame %>%
  select(1:5)%>%
  head(3)
losses_tidy <- losses_data_frame %>%
  rename(Team="...1")%>%
  pivot_longer(cols=-Team,names_to="Years",values_to = "Losses_of_Total")%>%
  separate(col="Losses_of_Total",into = c("Losses","Total"),sep = " of ", convert = TRUE)
hockey_df <- full_join(wins_tidy,losses_tidy)
hockey_df <- hockey_df%>%
  mutate(Draws=Total-Wins-Losses, across(.cols = -Total & is.numeric, ~.x/Total, .names ="{.col}_rt"))
head(hockey_df)

summary_hockey_df <- hockey_df %>%
  group_by(Team)%>%
  summarise(across(ends_with("_rt"),list(mean=mean, median=median)))
summary_hockey_df<- summary_hockey_df%>%
  rename(W_rt_mean=Wins_rt_mean, L_rt_mean=Losses_rt_mean,D_rt_mean=Draws_rt_mean, W_rt_median=Wins_rt_median, L_rt_meidan=Losses_rt_median,D_rt_median=Draws_rt_median)%>%
  arrange(desc(W_rt_median))
head(summary_hockey_df)
# L represents Lose, W represents Win and D represents Draw

```
### 2.3 Most correlated variables 
```{r}
max_cor_var<- function(df,col_name){
  v_col<-df%>%select(all_of(col_name))
  df_num<-df%>%
    select_if(is.numeric)%>%
    select(-all_of(col_name))
  correlations<-unlist(map(df_num,function(x){cor(x,v_col, use="complete.obs")}))
 max_abs_cor_var<-names(which(abs(correlations)==max(abs(correlations))))
 cor<-as.double(correlations[max_abs_cor_var])
 return(data.frame(var_name=max_abs_cor_var,cor=cor))
}

top_correlates_by_var <- function(df){
  num_df <- df%>% select(is.numeric)
  num_name <-names(num_df)
  corr_cols <- as.vector(unlist(map(num_name,function(.x){max_cor_var(df,.x)$var_name})))
  corr_cols_df<-data.frame(key=names(num_df),value=corr_cols)%>%
    pivot_wider(names_from = key,values_from = value)
  return(corr_cols_df)
}
penguins%>%top_correlates_by_var()


penguins%>%
  group_by(species)%>%
  nest()%>%
  mutate(max_var= map(data,top_correlates_by_var))%>%
  select(-data)%>%
  unnest(cols = max_var)



```



## 3 Elementary set theory  

### 3.1  Rolling a diceLet’s suppose we role a dice.  

   What is $A\cup{B}$? {1,2,3,4,6}     
   What is $A\cup{C}$? {1,2,3,4,5,6}     
   What is $A\cap{B}$? {2}  
   What is $A\cap{C}$? $\emptyset$  
   A\B = {1,3}  
   A\c = $\emptyset$  
   Are A and B disjoint? No  
   Are A and C disjoint? Yes  
   a partition of $\Omega$ = {1,2,3},{4,5,6} (2 sets)  {1,2}{3,4}{5,6} (3 sets)
### 3.2  Complements, subsets and De Morgan’s laws  
   Can you give an expression for$(A^c)^c$ without using the notion of a complement?
    $$(A^c)^c= A$$ 
    What is $\Omega^c$? $\emptyset$  
    Show that if A $A\subseteq{B}\ then B^c\subseteq{A^c}$   
    $Because\ B^c=\{\omega\in\Omega:\omega\notin{B}\}$  
    $And\ A^c=\{\omega\in\Omega:\omega\notin{A}\}$  
    $Then\ A\subseteq{B}$  
    $So\ B^c\subseteq{A^c}$  
   
### 3.3 Cardinality and the set of all subsets  
    $cardinality = 2^k$
### 3.4 isjointness and partitions   
    There is no set which is disjoint from any other sets
    $S_1,S_2,S_3,S_4 are pair-wise disjoint$
    $$S_1\cup{S_2}\cup{S_3}\cup{S_4}=A_1\cup{A_2}\cup{A_3}\cup{A_4}=\Omega$$
## 4 Introduction to probability


   