---
title: "Assignment 5"
author: "Ruinan Wang"
date: "27/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(purrr)
library(ggplot2)
```


### 2.Simulating data with the uniform distribution
```{r}
set.seed(0)
n<-1000
sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
    (0<=U)&(U<0.25)~1,
    (0.25<=U)&(U<0.5)~5,
  (0.5<=U)&(U<=1)~0))%>%
  pull(X)
```

#### why does this sample_X correspond to a sequence of i.i.d. copies X_1,....X__n of X where P(X=1) = a and P(X=5)=b and P(X=0) = 1-a-b with a=b=0.25?   
  runif function can generate the random numbers with uniform distribution in the range from 0 to 1. And set.seed() can generate the seed element of random number. case_when can get the number corresponds to the probability.
  

#### Now create a function called sample_X_015() which takes as inputs a,b and n and outputs a sample X_1,....X_n of indenpendent copies of X where P (X=1)=a and P (X=5)=b and P (X=0)=1-a-b.

```{r}
sample_x_015<-function (a,b,n){
  
  if (a>1||a<0||b>1||b<0||a+b>1||!is.integer(n))
    return("there is something wrong")
  else{
    sample_015<-data.frame(U = runif(n))%>%
      mutate(X=case_when(
        (0<=U)&(U<a)~1,
        (a<=U)&(U<a+b)~5,
        (a+b<=U)&(U<=1)~0))%>%
      pull(X)
    return(sample_015)
  }
}

n<-10000L
alpha<-1/2
beta<-1/10
sample_result <- sample_x_015(alpha,beta,n)
mean(sample_result)
expe <- 0.5*1+0.1*5
expe
```

when the number of sample becomes more and more, the mean average of sample is more close to the expectation of the sample

```{r}
vari <- var(sample_result)
vari
```



```{r}
set.seed(0)
sample_list <- data.frame(B_prob =seq(0,0.9,by=0.01))%>%
  mutate(samples_average = unlist(map(B_prob,~mean(sample_x_015(0.1,.x,100L)))))%>%
  mutate(expectation=0.1+B_prob*5)
ggplot(data=sample_list, aes(x=B_prob, y=samples_average))+geom_point()+geom_smooth(method="lm")
```



### 3 the Gaussian distribution

```{r}
# X <- seq(-4,6,length=100)
# y1 <- dnorm(X,1,1)
# y2 <- dnorm(X,1,sqrt(2))
# y3 <- dnorm(X,1,sqrt(3))
# dataSample <- data.frame(X,y1,y2,y3) 
# gaussianDF <- dataSample%>%
#   pivot_longer(starts_with("y"),names_to="Variance", values_to="Y" )
# head(gaussianDF)
# ggplot(data = gaussianDF, aes(x = X, y = Y, color=Type, linetype = Type))+ geom_line()

x<-seq(-4,6,0.1)
normal_densities_by_x<-data.frame(x=x,density=dnorm(x,mean=1,sd=sqrt(1)),var=1)%>%
  rbind(data.frame(x=x,density=dnorm(x,mean=1,sd=sqrt(2)),var=2))%>%
  rbind(data.frame(x=x,density=dnorm(x,mean=1,sd=sqrt(3)),var=3))
ggplot(normal_densities_by_x,aes(x=x,y=density, color=as.character(var), linetype = as.character(var))) + geom_line()+theme_bw()+labs(color="Variance",linetype = "Variance", x="x",y="Density")


cumulative_distribution_by_x<-data.frame(x=x,distribution=pnorm(x,mean=1,sd=sqrt(1)),var=1)%>%
  rbind(data.frame(x=x,distribution=pnorm(x,mean=1,sd=sqrt(2)),var=2))%>%
  rbind(data.frame(x=x,distribution=pnorm(x,mean=1,sd=sqrt(3)),var=3))

ggplot(cumulative_distribution_by_x,aes(x=x,y=distribution, color=as.character(var), linetype = as.character(var))) + geom_line()+theme_bw()+labs(color="Variance",linetype = "Variance", x="x",y="cumulative distribution function")



probs <- seq(0,1,0.01)
quantile_function_by_x<-data.frame(p=probs,quantile=qnorm(probs,mean=1,sd=sqrt(1)),var=1)%>%
  rbind(data.frame(p=probs,quantile=qnorm(probs,mean=1,sd=sqrt(2)),var=2))%>%
  rbind(data.frame(p=probs,quantile=qnorm(probs,mean=1,sd=sqrt(3)),var=3))

ggplot(quantile_function_by_x,aes(x=p,y=quantile, color=as.character(var), linetype = as.character(var))) + geom_line()+theme_bw()+labs(color="Variance",linetype = "Variance", x="x",y="quantile function")




set.seed(0)
standardGaussianSample <- rnorm(100)
mean1Var3GaussianSampleA <- standardGaussianSample * sqrt(3) + 1

set.seed(0)
mean1Var3GaussianSampleB <- rnorm(100, 1, sqrt(3))
all.equal(mean1Var3GaussianSampleA,mean1Var3GaussianSampleB)


colors<-c("Population density"="red", "Sample kernel density"="blue","Population mean"="green", "Sample mean"="violet")
linetypes<-c("Population density"="solid", "Sample kernel density"="dashed","Population mean"="solid", "Sample mean"="dashed")

ggplot()+labs(x="x",y="Density")+theme_bw()+geom_line(data= filter(normal_densities_by_x,var==3),aes(x,y=density,color="Population density")) +geom_density(data=data.frame(x=mean1Var3GaussianSampleA),aes(x=x,color="Sample kernel density", linetype = "Sample kernel density")) +geom_vline(aes(xintercept=1,color="Population mean", linetype = "Population mean")) + geom_vline(aes(xintercept=mean(mean1Var3GaussianSampleA),color="Sample mean", linetype="Sample mean")) + scale_color_manual(name="Legend",values=colors)+scale_linetype_manual(name="Legend",values=linetypes)
```


### 4 The Binomial distribution and the central limit theorem

```{r}
p<- 0.7
n<- 50

binom_df<- data.frame(x=seq(0,n))%>%
  mutate(pmf=map_dbl(.x=x,~dbinom(.x,size=n,prob=p)))
head(binom_df)

mu = n*p
sigma = sqrt(n*p*(1-p))
gaussian_df <- data.frame(x=seq(0,n, by=0.01))%>%
  mutate(pdf=map_dbl(.x=x,~dnorm(.x,mean=mu,sd=sigma)))
head(gaussian_df)

colors<- c("Gaussian pdf"="red","Binomial pmf"="blue")
fill <- c("Gaussian pdf"="white", "Binomial pmf"="white")

ggplot()+labs(x="x",y="Probability")+theme_bw()+
  geom_line(data=gaussian_df,aes(x=x,y=pdf, color="Gaussian pdf"), size=2) + geom_col(data=binom_df, aes(x=x,y=pmf, color="Binomial pmf", fill="Binomial pmf")) + scale_color_manual(name="",values=colors)+scale_fill_manual(name="",values=fill)+xlim(20,50)
```



### 5 Exponential distribution

```{r}
my_cdf_exp <- function (x,lambda) {
  if (x<=0){
    return(0)
  }
  else {
    return (1-exp(-lambda*x))
  }
} 
  
lambda <- 1/2

map_dbl(.x=seq(-1,4),.f=~my_cdf_exp(x=.x,lambda=lambda))


test_inputs<-seq(-1,10,0.1)
my_cdf_output<-map_dbl(.x=test_inputs, .f=~my_cdf_exp(x=.x,lambda=lambda))
inbuilt_cdf_output<-map_dbl(.x=test_inputs, .f=~pexp(q=.x,rate=lambda))
all.equal(my_cdf_output,inbuilt_cdf_output)

my_quantile_exp <- function (p,lambda){
  return(-1/lambda*log(1-p,exp(1)))
}

p<- seq(0.01,0.99,0.01)
my_quantile_output<-map_dbl(.x=p,.f=~my_quantile_exp(.x,lambda))

inbuilt_quantile_output<- map_dbl(p,~qexp(.x,lambda))

all.equal(my_quantile_output,inbuilt_quantile_output)

```



